---
title: "N-Mixture/CMR Exploratory - Pennsylvania"
format: pdf
---

## Load Packages

```{r}
#| output: false
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(tidyverse, RPostgres, DBI, unmarked, here, lubridate, kableExtra)

```

## Connect to `survey_data` schema in `ribbitr` database
```{r}
#| output: false
tryCatch({
    drv <- dbDriver("Postgres")
    print("Connecting to Database…")
    connection <- dbConnect(drv,
                 dbname = Sys.getenv("aws_dbname"),
                 host = Sys.getenv("aws_host"),
                 port = Sys.getenv("aws_port"),
                 user = Sys.getenv("aws_user"),
                 password = Sys.getenv("aws_password"),
                 timezone=NULL)
    print("Database Connected!")
    },
    error=function(cond) {
            print("Unable to connect to Database.")
    })

#search path
dbExecute(connection, "set search_path to survey_data")



```

## Query 2022 CMR Penn data
```{r}

# Data
cmr_q <- "select r.region, s.site, v.date, v.survey_time, s2.duration_minutes, 
          c.species_capture, c.capture_type
          from region r
          join site s on r.region_id = s.region_id 
          full join visit v on s.site_id = v.site_id 
          join survey s2 on v.visit_id = s2.visit_id 
          join capture c on s2.survey_id = c.survey_id
          where r.region = 'pennsylvania'
          and v.date > '2022-01-01';"

raw_data <- dbGetQuery(connection, cmr_q) %>% 
  select(!c(region, survey_time, duration_minutes)) %>% 
  arrange(date)



# find all visits
visit_cmr_q <- "select r.region, s.site, v.date, v.survey_time
                from region r
                join site s on r.region_id = s.region_id 
                join visit v on s.site_id = v.site_id 
                where r.region = 'pennsylvania'
                and v.date > '2022-01-01';"

raw_visits <-dbGetQuery(connection, visit_cmr_q) %>% 
  arrange(date) %>% 
  select(site, date)




```

## Visualize all Visits
```{r}
viz <- raw_visits %>% 
  group_by(site) %>% 
  summarise(n = n())

ggplot(data = viz) +
  geom_col(aes(x=site, y = n)) +
  ggtitle("Raw Visits")

```


## Clean up species list with counts
#### Filter for CMR focal species and summarize species counts. Then populate zeros for 3 focal species into the data set for visits when captures did not occur.
```{r}
#| output: false

# filter for CMR focal species and summarize counts
mid_clean_up <- raw_data %>% 
  filter(species_capture %in% c("pseudacris_crucifer", "rana_catesbeiana", 
                                "rana_clamitans")) %>% 
  mutate(capture_type = if_else(is.na(capture_type), "new", capture_type)) %>% 
  group_by(date, site, species_capture) %>% 
  summarise(n = n()) %>% 
  ungroup() 

# populate zeros
clean_up <- raw_visits %>% 
  left_join(mid_clean_up) %>%
  complete(nesting(date, site), 
           species_capture = unique(mid_clean_up$species_capture), 
           fill = list(n = 0))


```

## rana_catesbeiana

### Table formatting
#### Filter data for `rana_catesbeiana`, tally the numbner of visits, pivot data frame into correct matrix form, and finally populate zeros into the NAs if sites were visited. 
```{r}

bull_frog_visits <- clean_up %>% 
  select(site, date, species_capture, n) %>% 
  filter(species_capture == "rana_catesbeiana") %>% 
  select(!species_capture) %>% 
  group_by(site) %>% 
  mutate(n_visit = match(date, unique(date)),
         n_visit = paste0("visit_", n_visit, sep = "")) %>% 
  select(!date) %>% 
  ungroup() %>%
  pivot_wider(names_from = c("n_visit"), values_from = c("n")) %>% 
  group_by(site) %>% 
  mutate(across(contains("visit"), 
                ~ifelse(is.na(.) &
                          !is.na(lag(.)), 0, .)))



kableExtra::kable(bull_frog_visits)

```

### static n-mixture models no covariates 
```{r}

bull_frog_unmarked_nmixture <- bull_frog_visits %>% 
  ungroup() %>% 
  select(!c(site)) 


umf <- unmarkedFramePCount(y = bull_frog_unmarked_nmixture)

summary(umf)

fm1 <- pcount(~1 ~1, data = umf)

# Get estimates on link scale 
backTransform(fm1, "state") 
# Get estimates on natural scale
backTransform(fm1, "det")

```
#### Assuming the sites and enviro variables are exactly the same we can say there is 40.7 +-7.9 frogs at these sites and we are only detecting 23% of the frogs.


## pseudacris_crucifer
### Table formatting
#### Filter data for `pseudacris_crucifer`, tally the numbner of visits, pivot data frame into correct matrix form, and finally populate zeros into the NAs if sites were visited.
```{r}

peep_frog_visits <- clean_up %>% 
  select(site, date, species_capture, n) %>% 
  filter(species_capture == "pseudacris_crucifer") %>% 
  select(!species_capture) %>% 
  group_by(site) %>% 
  mutate(n_visit = match(date, unique(date)),
         n_visit = paste0("visit_", n_visit, sep = "")) %>% 
  select(!date) %>% 
  ungroup() %>%
  pivot_wider(names_from = c("n_visit"), values_from = c("n"))%>% 
  group_by(site) %>% 
  mutate(across(contains("visit"), 
                ~ifelse(is.na(.) &
                          !is.na(lag(.)), 0, .)))


kable(peep_frog_visits)

```

### static n-mixture unmakred models no covariates

```{r}
peep_unmarked_nmixture <- peep_frog_visits %>% 
  ungroup() %>% 
  #filter(capture_type == "new") %>% 
  select(!c(site)) 


umf <- unmarkedFramePCount(y = peep_unmarked_nmixture)

summary(umf)

fm1 <- pcount(~1 ~1, data = umf)

backTransform(fm1, "state") 

backTransform(fm1, "det")



```
#### Assuming the sites and enviro variables are exactly the same we can estimate there is 85.7 +-24.4 frogs at these sites and we are only detecting 1.2% of the frogs.


## rana_clamitans

### Table formatting
#### Filter data for `rana_clamitans`, tally the numbner of visits, pivot data frame into correct matrix form, and finally populate zeros into the NAs if sites were visited.

```{r}

green_frog_visits <- clean_up %>% 
  select(site, date, species_capture, n) %>% 
  filter(species_capture == "rana_clamitans") %>% 
  #select(!species_capture) %>% 
  group_by(site) %>% 
  mutate(n_visit = match(date, unique(date)),
         n_visit = paste0("visit_", n_visit, sep = "")) %>% 
  select(!date) %>% 
  ungroup() %>%
  group_by(site, n_visit, ) %>% 
  summarise(n = sum(n)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = c("n_visit"), values_from = c("n")) %>% 
  #add_row(site = "phelps_pond", capture_type = "recapture") %>% 
  group_by(site) %>% 
  mutate(across(contains("visit"), 
                ~ifelse(is.na(.) &
                          !is.na(lag(.)), 0, .))) 



```

### static n-mixture with no co-variates

```{r}

green_frog_unmarked_nmixture <- green_frog_visits %>% 
  ungroup() %>% 
  select(!c(site)) 


umf <- unmarkedFramePCount(y = green_frog_unmarked_nmixture)

summary(umf)

fm1 <- pcount(~1 ~1, data = umf)

backTransform(fm1, "state") 

backTransform(fm1, "det")



```
#### Assuming the sites and enviro variables are exactly the same we estimate there is 94.8 +-13.1 frogs at these sites and we are only detecting 4.3% of the frogs.


## Table formating capture-recapture

```{r}


# mid_clean_up <- raw_data %>% 
#   filter(species_capture %in% c("pseudacris_crucifer", "rana_catesbeiana", 
#                                 "rana_clamitans")) %>% 
#   mutate(capture_type = if_else(is.na(capture_type), "new", capture_type)) %>% 
#   group_by(date, site, capture_type, species_capture) %>% 
#   summarise(n = n()) %>% 
#   ungroup() 
# 
# # populate zeros
# clean_up <- raw_visits %>% 
#   left_join(mid_clean_up) %>%
#   complete(nesting(date, site), 
#            species_capture = unique(mid_clean_up$species_capture), 
#            fill = list(n = 0)) %>% 
#   mutate(capture_type = if_else(is.na(capture_type), "new", capture_type))
# 


```

### Bullfrog CMR data

```{r}

# bull_frog_cmr <- clean_up %>% 
#   select(site, date, capture_type, species_capture, n) %>% 
#   filter(species_capture == "rana_catesbeiana") %>% 
#   select(!species_capture) %>% 
#   group_by(site) %>% 
#   mutate(n_visit = match(date, unique(date)),
#          n_visit = paste0("visit_", n_visit, sep = "")) %>% 
#   select(!date) %>% 
#   ungroup() %>%
#   pivot_wider(names_from = c("n_visit"), values_from = c("n")) %>% 
#   add_row(site = "phelps_pond", capture_type = "recapture") %>%
#   add_row(site = "tryon_weber", capture_type = "recapture", visit_1 = 0, visit_2 = 0, visit_3 = 0) %>% 
#   group_by(site) %>% 
#   mutate(across(contains("visit"), 
#                 ~ifelse(is.na(.) &
#                           !is.na(lag(.)), 0, .)))

```

```{r}

# t <- clean_up %>% 
#   select(site, date, species_capture, n) %>% 
#   filter(species_capture == "rana_catesbeiana") %>% 
#   select(!species_capture) %>% 
#   group_by(site) %>% 
#   mutate(n_visit = match(date, unique(date)),
#          n_visit = paste0("visit_", n_visit, sep = "")) %>% 
#   select(!date) %>% 
#   ungroup() %>%
#   pivot_wider(names_from = c("n_visit"), values_from = c("n")) %>% 
#   group_by(site) %>% 
#   mutate(across(contains("visit"), 
#                 ~ifelse(is.na(.) &
#                           !is.na(lag(.)), 0, .)))



```
