---
title: "Exploratory N-Mixture/CMR - Pennsylvania"
format: pdf
---

## Load Packages

```{r}
#| output: false
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(tidyverse, RPostgres, DBI, unmarked, here, lubridate, kableExtra)

```

## Connect to `survey_data` schema in `ribbitr` database
```{r}
#| output: false
tryCatch({
    drv <- dbDriver("Postgres")
    print("Connecting to Database…")
    connection <- dbConnect(drv,
                 dbname = Sys.getenv("aws_dbname"),
                 host = Sys.getenv("aws_host"),
                 port = Sys.getenv("aws_port"),
                 user = Sys.getenv("aws_user"),
                 password = Sys.getenv("aws_password"),
                 timezone=NULL)
    print("Database Connected!")
    },
    error=function(cond) {
            print("Unable to connect to Database.")
    })

#search path
dbExecute(connection, "set search_path to survey_data")



```

## Query 2022 N-Mix Penn data
```{r}

# Data
nmix_q <- "select r.region, s.site, v.date, v.survey_time, s2.duration_minutes, 
          v2.species_ves, v2.count
          from region r
          join site s on r.region_id = s.region_id 
          full join visit v on s.site_id = v.site_id 
          join survey s2 on v.visit_id = s2.visit_id 
          join ves v2 on s2.survey_id = v2.survey_id
          where r.region = 'pennsylvania'
          and v.date > '2023-01-01';"

nmix_raw_data <- dbGetQuery(connection, nmix_q) %>% 
  select(!c(region, survey_time, duration_minutes)) %>% 
  arrange(date) 

#write_csv(nmix_raw_data, here("data", "nmix_raw_data.csv"))



# find all visits
visit_nmix_q <- "select r.region, s.site, v.date, v.survey_time, s2.detection_type, s2.observer
                from region r
                join site s on r.region_id = s.region_id 
                join visit v on s.site_id = v.site_id 
                join survey s2 on v.visit_id = s2.visit_id
                where r.region = 'pennsylvania'
                and v.date > '2023-01-01'
                and s2.detection_type = 'visual';"

nmix_raw_visits <-dbGetQuery(connection, visit_nmix_q) %>% 
  arrange(date) %>% 
  select(site, date)


#write_csv(nmix_raw_visits, here("data", "nmix_raw_visits.csv"))

```

## Visualize all VES Visits
```{r}
#nmix_raw_visits <- read_csv(here("data", "nmix_raw_visits.csv"))
#nmix_raw_data <- read_csv(here("data", "nmix_raw_data.csv"))


viz <- nmix_raw_visits %>% 
  group_by(site) %>% 
  summarise(n = n())

ggplot(data = viz) +
  geom_col(aes(x=site, y = n)) +
  ggtitle("Raw Visits")

```
\newpage

## Clean up species list with counts
#### Filter for CMR focal species and summarize species counts. Then populate zeros for 3 focal species into the data set for visits when captures did not occur.
```{r}
#| output: false

# filter for CMR focal species and summarize counts
n_mix_mid_clean_up <- nmix_raw_data %>% 
  filter(species_ves %in% c("pseudacris_crucifer", "rana_catesbeiana", 
                                "rana_clamitans", "rana_pipiens")) %>% 
  #mutate(capture_type = if_else(is.na(capture_type), "new", capture_type)) %>% 
  group_by(date, site, species_ves) %>% 
  summarise(n = sum(count)) %>% 
  ungroup() 

# populate zeros
nmix_clean_up <- n_mix_mid_clean_up %>% 
  #left_join(n_mix_mid_clean_up) %>%
  complete(nesting(date, site), 
           species_ves = unique(n_mix_mid_clean_up$species_ves), 
           fill = list(n = 0))




```

```{r}
kable(head(nmix_clean_up, n = 15))
```

\newpage

## rana_catesbeiana

### N-Mixture Table formatting
#### Filter data for `rana_catesbeiana`, tally the numbner of visits, pivot data frame into correct matrix form, and finally populate zeros into the NAs if sites were visited. 
```{r}

bull_frog_visits <- nmix_clean_up %>% 
  select(site, date, species_ves, n) %>% 
  filter(species_ves == "rana_catesbeiana") %>% 
  select(!species_ves) %>% 
  group_by(site) %>% 
  mutate(n_visit = match(date, unique(date)),
         n_visit = paste0("visit_", n_visit, sep = "")) %>% 
  select(!date) %>% 
  ungroup() %>%
  pivot_wider(names_from = c("n_visit"), values_from = c("n")) %>% 
  group_by(site) %>% 
  mutate(across(contains("visit"), 
                ~ifelse(is.na(.) &
                          !is.na(lag(.)), 0, .)))



kable(bull_frog_visits)

```

### Bullfrog static n-mixture models no covariates - 3 repeat sites
```{r}

first_bull_frog_unmarked_nmixture <- bull_frog_visits %>% 
  ungroup() %>% 
  select(!c(site, c(visit_4:visit_6))) %>% 
  slice(2:4)

second_bull_frog_unmarked_nxmixture <- bull_frog_visits %>% 
  ungroup() %>% 
  select(site, c(visit_4:visit_6)) %>% 
  slice(2) %>% 
  rename(visit_1 = visit_4,
         visit_2 = visit_5,
         visit_3 = visit_6) %>% 
  select(!c(site))

bull_frog_nmixture_3_visits <- rbind(first_bull_frog_unmarked_nmixture, second_bull_frog_unmarked_nxmixture)

bf_umf_3 <- unmarkedFramePCount(y = bull_frog_nmixture_3_visits)

summary(bf_umf_3)

fm1_bf3 <- pcount( ~ 1 ~ 1, 
               data = bf_umf_3, 
               engine = "R")

summary(fm1_bf3)

backTransform(fm1_bf3, "state")

backTransform(fm1_bf3, "det")


```
#### Assuming the sites and enviro variables are exactly the same we can say there is 18.08 +-4.48 frogs at each site with 65 % chance of detecting each individual when 3 repeat visits occur nightly.

\newpage

### Bullfrog static n-mixture models no covariates - 2 repeat sites
```{r}

first_bull_frog_unmarked_nmixture <- bull_frog_visits %>% 
  ungroup() %>% 
  select(!c(site, c(visit_3:visit_6))) %>% 
  slice(2:4)

second_bull_frog_unmarked_nxmixture <- bull_frog_visits %>% 
  ungroup() %>% 
  select(site, c(visit_4:visit_5)) %>% 
  slice(2) %>% 
  rename(visit_1 = visit_4,
         visit_2 = visit_5) %>% 
  select(!c(site))

bull_frog_nmixture_2_visits <- rbind(first_bull_frog_unmarked_nmixture, second_bull_frog_unmarked_nxmixture)

bf_umf_2 <- unmarkedFramePCount(y = bull_frog_nmixture_2_visits)

summary(bf_umf_2)

fm1_bf2 <- pcount( ~ 1 ~ 1, 
               data = bf_umf_2, 
               engine = "R")

summary(fm1_bf2)

backTransform(fm1_bf2, "state")

backTransform(fm1_bf2, "det")


```

#### Assuming the sites and enviro variables are exactly the same we can say there is 13.91 +-3.18 frogs at each site with 79 % chance of detecting each individual when 2 repeat visits occur nightly.

\newpage



## rana_clamitans

### N-Mixture Table Formatting
#### Filter data for `rana_clamitans`, tally the numbner of visits, pivot data frame into correct matrix form, and finally populate zeros into the NAs if sites were visited.

```{r}

green_frog_visits <- nmix_clean_up %>% 
  select(site, date, species_ves, n) %>% 
  filter(species_ves == "rana_clamitans") %>% 
  #select(!species_capture) %>% 
  group_by(site) %>% 
  mutate(n_visit = match(date, unique(date)),
         n_visit = paste0("visit_", n_visit, sep = "")) %>% 
  select(!date) %>% 
  ungroup() %>%
  group_by(site, n_visit, ) %>% 
  summarise(n = sum(n)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = c("n_visit"), values_from = c("n")) %>% 
  #add_row(site = "phelps_pond", capture_type = "recapture") %>% 
  group_by(site) %>% 
  mutate(across(contains("visit"), 
                ~ifelse(is.na(.) &
                          !is.na(lag(.)), 0, .))) 


kable(green_frog_visits)

```

### static n-mixture with no co-variates - 3 repeat visits

```{r}

first_green_frog_unmarked_nmixture <- green_frog_visits %>%  
  ungroup() %>% 
  select(!c(site, c(visit_4:visit_6))) %>% 
  slice(1:3) 

second_green_frog_unmarked_nxmixture <- green_frog_visits %>% 
 ungroup() %>% 
  select(site, c(visit_4:visit_6)) %>% 
  slice(1) %>% 
  rename(visit_1 = visit_4,
         visit_2 = visit_5,
         visit_3 = visit_6) %>% 
  select(!c(site))

green_frog_nmixture_3_visits <- rbind(first_green_frog_unmarked_nmixture, second_green_frog_unmarked_nxmixture)


umf_green3 <- unmarkedFramePCount(y = green_frog_nmixture_3_visits)

summary(umf_green3)

fm1_g3 <- pcount(~1 ~1, data = umf_green3)

backTransform(fm1_g3, "state") 

backTransform(fm1_g3, "det")



```
$~$

#### Assuming the sites and enviro variables are exactly the same we estimate there is 25.42 +-9.83 frogs at each sites and with 31% of detecting each  individual when 3 repeat visits occur nightly.

\newpage

### static n-mixture with no co-variates - 2 repeat visits

```{r}

first_green_frog_unmarked_nmixture <- green_frog_visits %>% 
  ungroup() %>% 
  select(!c(site, c(visit_3:visit_6))) %>% 
  slice(1:3) 

second_green_frog_unmarked_nxmixture <- green_frog_visits %>% 
 ungroup() %>% 
  select(site, c(visit_4:visit_5)) %>% 
  slice(1) %>% 
  rename(visit_1 = visit_4,
         visit_2 = visit_5) %>% 
  select(!c(site))

green_frog_nmixture_2_visits <- rbind(first_green_frog_unmarked_nmixture, second_green_frog_unmarked_nxmixture)


umf_green2 <- unmarkedFramePCount(y = green_frog_nmixture_2_visits)

summary(umf_green2)

fm1_g2 <- pcount(~1 ~1, data = umf_green2)

backTransform(fm1_g2, "state") 

backTransform(fm1_g2, "det")



```
$~$

#### Assuming the sites and enviro variables are exactly the same we estimate there is 25.42 +-9.83 frogs at each sites and with 31% of detecting each  individual when 3 repeat visits occur nightly.

\newpage



## Query CMR data
```{r}
# Data
cmr_q <- "select r.region, s.site, v.date, v.survey_time, s2.duration_minutes, 
          c.species_capture, c.capture_type, cmr.cmr_id
          from region r
          join site s on r.region_id = s.region_id 
          full join visit v on s.site_id = v.site_id 
          join survey s2 on v.visit_id = s2.visit_id 
          join capture c on s2.survey_id = c.survey_id
          join penn_cmr cmr on c.capture_mark_recapture = cmr.capture_mark_recapture
          where r.region = 'pennsylvania'
          and v.date > '2023-01-01';"

cmr_raw_data <- dbGetQuery(connection, cmr_q) %>% 
  select(!c(region, survey_time, duration_minutes)) %>% 
  arrange(date)

#write_csv(cmr_raw_data, here("data", "cmr_raw_data.csv"))

# find all visits
visit_cmr_q <- "select r.region, s.site, v.date, v.survey_time, s2.detection_type
                from region r
                join site s on r.region_id = s.region_id 
                join visit v on s.site_id = v.site_id 
                join survey s2 on v.visit_id = s2.visit_id
                where r.region = 'pennsylvania'
                and v.date > '2023-01-01'
                and s2.detection_type = 'capture';"

cmr_raw_visits <-dbGetQuery(connection, visit_cmr_q) %>% 
  arrange(date) %>% 
  select(site, date)

#write_csv(cmr_raw_visits, here("data", "cmr_raw_visits.csv"))
```

## Bullfrog CMR data
### Matrix: individual counts by visits
```{r}

#cmr_raw_data <- read_csv(here("data", "cmr_raw_data.csv"))
#mr_raw_visits <- read_csv(here("data", "cmr_raw_visits.csv"))

bull_mid_clean_up <- cmr_raw_data %>% 
  filter(species_capture == "rana_catesbeiana") %>% 
  select(!c(capture_type)) %>% 
  unite(species_capture, c(species_capture, cmr_id), sep = "_")
  
bull_pop_zeros <- cmr_raw_visits %>% 
  left_join(bull_mid_clean_up) %>%
  complete(nesting(date, site), 
           fill = list(n = 0)) %>% 
  ungroup() %>% 
  group_by(site) %>% 
  mutate(n_visit = match(date, unique(date)),
         n_visit = paste0("visit_", n_visit, sep = "")) %>% 
  select(!date) %>% 
  ungroup() 
  
clean_bull <- bull_pop_zeros %>%   
  group_by(site, species_capture, n_visit) %>% 
  reframe(n = n()) %>% 
  mutate(n = if_else(is.na(species_capture), NA, n)) %>% 
  drop_na(species_capture) %>%
  pivot_wider(names_from = c("n_visit"), values_from = c("n")) %>%
  #select(!c(visit_5, visit_6)) %>% 
  # relocate(visit_1, .before = visit_2) %>% 
  # relocate(visit_3, .after = visit_2) %>% 
  
  # mutate(across(contains("visit"), 
  #               ~ifelse(is.na(.) &
  #                         !is.na(lag(.)), 0, .))) %>% 
  #select(sort(names(.))) %>% 
  #filter(!site == "tryon_weber") %>% 
  select(!site) %>% 
  mutate_all(~replace_na(.,0)) 
  
  
```

### Final CMR matrix
```{r}

clean_bull$captureHistory <- paste(clean_bull$visit_1, clean_bull$visit_2, clean_bull$visit_3,
                               sep = "")

# clean_bull <- clean_bull %>% 
#   ungroup() %>% 
#   group_by(species_capture, captureHistory) %>% 
#   unique()

lev <- unique(clean_bull$captureHistory)

clean_bull$captureHistory <- factor(clean_bull$captureHistory, levels = lev)

bull_table <- table(clean_bull$species_capture, clean_bull$captureHistory)
  

kable(head(bull_table, n = 15))

```

### obs covariates, piFun equal detection, Model runn
```{r}
class(bull_table) <- "matrix"

o2y <- matrix(1, 3, 6)

crPiFun <- function(p) {
    p1 <- p[,1]
    p2 <- p[,2]
    p3 <- p[,3]

    cbind("100" = p1 * (1-p2) * (1-p3),    #1: 100    
          "010" = p1 * (1-p2) * (1-p3),    #2: 010
          "001" = (1-p1) * (1-p2) * p3,    #3: 001
          "111" = p1 * p2 * p3,            #4: 111
          "110" = p1 * p2 * (1-p3),        #5: 110
          "011" = (1-p1) * p2 * p3         #6: 011
          
    )
}


umf <- unmarkedFrameMPois(y = bull_table, piFun = "crPiFun", obsToY = o2y)

fm <- multinomPois(~1 ~1, umf, engine = "R")

backTransform(fm, "state")
backTransform(fm, "det")
```

#### Assuming sites are identical we estimate there are 8 individuals at each site? With a 30% chance of detecting each individual?  Not fully sure how to interpret results. 

\newpage

# STOP: Species Not detected in 2023 data for VES

## pseudacris_crucifer
### N-Mixture Table formatting
#### Filter data for `pseudacris_crucifer`, tally the numbner of visits, pivot data frame into correct matrix form, and finally populate zeros into the NAs if sites were visited.
```{r}

peep_frog_visits <- nmix_clean_up %>% 
  select(site, date, species_ves, n) %>% 
  filter(species_ves == "pseudacris_crucifer") %>% 
  select(!species_ves) %>% 
  group_by(site) %>% 
  mutate(n_visit = match(date, unique(date)),
         n_visit = paste0("visit_", n_visit, sep = "")) %>% 
  select(!date) %>% 
  ungroup() %>%
  pivot_wider(names_from = c("n_visit"), values_from = c("n"))%>% 
  group_by(site) %>% 
  mutate(across(contains("visit"), 
                ~ifelse(is.na(.) &
                          !is.na(lag(.)), 0, .)))


kable(peep_frog_visits)

```

### static n-mixture model no covariates
```{r}
peep_unmarked_nmixture <- peep_frog_visits %>% 
  ungroup() %>% 
  #filter(capture_type == "new") %>% 
  select(!c(site)) 


umf <- unmarkedFramePCount(y = peep_unmarked_nmixture)

summary(umf)

fm1 <- pcount(~1 ~1, data = umf)

backTransform(fm1, "state") 

backTransform(fm1, "det")



```
#### Assuming the sites and enviro variables are exactly the same we can estimate there is 85.7 +-24.4 frogs at each site with a 1.2% chance of detecting each individual.

\newpage

## rana_pipens

### N-Mixture Table Formatting
#### Filter data for `rana_pipens`, tally the numbner of visits, pivot data frame into correct matrix form, and finally populate zeros into the NAs if sites were visited.

```{r}

leop_frog_visits <- nmix_clean_up %>% 
  select(site, date, species_ves, n) %>% 
  filter(species_ves == "rana_pipiens") %>% 
  #select(!species_capture) %>% 
  group_by(site) %>% 
  mutate(n_visit = match(date, unique(date)),
         n_visit = paste0("visit_", n_visit, sep = "")) %>% 
  select(!date) %>% 
  ungroup() %>%
  group_by(site, n_visit, ) %>% 
  summarise(n = sum(n)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = c("n_visit"), values_from = c("n")) %>% 
  #add_row(site = "phelps_pond", capture_type = "recapture") %>% 
  group_by(site) %>% 
  mutate(across(contains("visit"), 
                ~ifelse(is.na(.) &
                          !is.na(lag(.)), 0, .))) 


kable(leop_frog_visits)

```

### static n-mixture with no co-variates

```{r}

leop_frog_unmarked_nmixture <- leop_frog_visits %>% 
  ungroup() %>% 
  select(!c(site)) 


umf <- unmarkedFramePCount(y = leop_frog_unmarked_nmixture)

summary(umf)

fm1 <- pcount(~1 ~1, data = umf)

backTransform(fm1, "state") 

backTransform(fm1, "det")



```

\newpage


