---
title: "N-Mix_Samples"
format: html
editor: visual
---

## Load Packages

```{r}
#| output: false
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(tidyverse, RPostgres, DBI, unmarked, here, lubridate, kableExtra, GGally, MuMIn, AHMbook, AICcmodavg)

```

## Connect to `survey_data` schema in `ribbitr` database
```{r}
#| output: false
tryCatch({
    drv <- dbDriver("Postgres")
    print("Connecting to Database…")
    connection <- dbConnect(drv,
                 dbname = Sys.getenv("aws_dbname"),
                 host = Sys.getenv("aws_host"),
                 port = Sys.getenv("aws_port"),
                 user = Sys.getenv("aws_user"),
                 password = Sys.getenv("aws_password"),
                 timezone=NULL)
    print("Database Connected!")
    },
    error=function(cond) {
            print("Unable to connect to Database.")
    })

#search path
dbExecute(connection, "set search_path to survey_data")



```

## Query 2022 N-Mix Penn data
```{r}

# Data
nmix_q <- "select r.region, s.site, s.area_sqr_m, v.date, v.survey_time, s2.duration_minutes, 
          v2.species_ves, v2.count, v2.observer, s2.air_temp_c, s2.water_temp_c, s2.wind, s2.sky, s2.relative_humidity_percent
          from region r
          join site s on r.region_id = s.region_id 
          full join visit v on s.site_id = v.site_id 
          join survey s2 on v.visit_id = s2.visit_id 
          join ves v2 on s2.survey_id = v2.survey_id
          where r.region = 'pennsylvania'
          and v.date > '2023-01-01';"

nmix_raw_data <- dbGetQuery(connection, nmix_q) %>% 
  select(!c(region, survey_time)) %>% 
  arrange(date) 

#write_csv(nmix_raw_data, here("data", "nmix_raw_data.csv"))



# find all visits
visit_nmix_q <- "select r.region, s.site, v.date, v.survey_time, s2.detection_type, s2.observer
                from region r
                join site s on r.region_id = s.region_id 
                join visit v on s.site_id = v.site_id 
                join survey s2 on v.visit_id = s2.visit_id
                where r.region = 'pennsylvania'
                and v.date > '2023-01-01'
                and s2.detection_type = 'visual';"

nmix_raw_visits <-dbGetQuery(connection, visit_nmix_q) %>% 
  arrange(date) %>% 
  select(site, date)


#write_csv(nmix_raw_visits, here("data", "nmix_raw_visits.csv"))

```

## Visualize all VES Visits
```{r}
#nmix_raw_visits <- read_csv(here("data", "nmix_raw_visits.csv"))
#nmix_raw_data <- read_csv(here("data", "nmix_raw_data.csv"))


viz <- nmix_raw_visits %>% 
  group_by(site) %>% 
  summarise(n = n())

ggplot(data = viz) +
  geom_col(aes(x=site, y = n)) +
  ggtitle("Raw Visits")

```
\newpage

## Clean up species list with counts
#### Filter for CMR focal species and summarize species counts. Then populate zeros for 3 focal species into the data set for visits when captures did not occur.
```{r}
#| output: false

# filter for CMR focal species and summarize counts
n_mix_mid_clean_up <- nmix_raw_data %>% 
  filter(species_ves %in% c("pseudacris_crucifer", "rana_catesbeiana", 
                                "rana_clamitans", "rana_pipiens")) %>% 
  #mutate(capture_type = if_else(is.na(capture_type), "new", capture_type)) %>% 
  group_by(date, site, species_ves) %>% 
  summarise(n = sum(count)) %>% 
  ungroup() 

# populate zeros
nmix_clean_up <- n_mix_mid_clean_up %>% 
  full_join(nmix_raw_visits) %>%
  complete(nesting(date, site), 
           species_ves = unique(n_mix_mid_clean_up$species_ves), 
           fill = list(n = 0)) %>% 
  drop_na(species_ves) %>% 
  filter(site == "rv_pond") # change for site




```


\newpage

## rana_catesbeiana

### N-Mixture Table formatting
#### Filter data for `rana_catesbeiana`, tally the numbner of visits, pivot data frame into correct matrix form, and finally populate zeros into the NAs if sites were visited. 
```{r}

bull_frog_visits <- nmix_clean_up %>% 
  select(site, date, species_ves, n) %>% 
  filter(species_ves == "rana_clamitans") %>% 
  select(!species_ves) %>% 
  group_by(site) %>% 
  mutate(n_visit = match(date, unique(date)),
         n_visit = paste0("visit_", n_visit, sep = "")) %>% 
  select(!date) %>% 
  ungroup() %>%
  pivot_wider(names_from = c("n_visit"), values_from = c("n")) %>% 
  group_by(site) %>% 
  mutate(across(contains("visit"), 
                ~ifelse(is.na(.) &
                          !is.na(lag(.)), 0, .)))




```



### Bullfrog static n-mixture models no covariates - 3 repeat sites
```{r}

bull_frog_nmixture_3_visits <- bull_frog_visits %>% 
  ungroup() %>% 
  select(!c(site))

write_csv(bull_frog_nmixture_3_visits, here("tuttle_bullfrog.csv"))

library(unmarked)

# 3 Visits
bull_frog_nmixture_3_visits <- read.csv("tuttle_bullfrog.csv")


bf_umf_3 <- unmarkedFramePCount(y = bull_frog_nmixture_3_visits)

summary(bf_umf_3)

fm1_bf3 <- pcount( ~ 1 ~ 1, 
               data = bf_umf_3, 
               engine = "R",
               control=list(trace=T, REPORT=5))

summary(fm1_bf3)


backTransform(fm1_bf3, "state")

backTransform(fm1_bf3, type = "det")


# 2 Visits
bull_frog_nmixture_2_visits <- bull_frog_nmixture_3_visits %>% 
  select(!c(visit_3))

bf_umf_2 <- unmarkedFramePCount(y = bull_frog_nmixture_2_visits)

summary(bf_umf_2)

fm1_bf2 <- pcount( ~ 1 ~ 1, 
               data = bf_umf_2, 
               engine = "R",
               control=list(trace=T, REPORT=5))

summary(fm1_bf2)


backTransform(fm1_bf2, "state")

backTransform(fm1_bf2, type = "det")
  


```

### Bull Frog RV

```{r}

first_bull_frog_nmixture_3_visits <- bull_frog_visits %>% 
  ungroup() %>% 
  select(!c(site, c(visit_4:visit_6)))

second_bull_frog_unmarked_nxmixture <- bull_frog_visits %>% 
  ungroup() %>% 
  select(site, c(visit_4:visit_6)) %>% 
  rename(visit_1 = visit_4,
         visit_2 = visit_5,
         visit_3 = visit_6) %>% 
  ungroup() %>% 
  select(!c(site))

# bull_frog_nmixture_3_visits <- rbind(first_bull_frog_nmixture_3_visits, second_bull_frog_unmarked_nxmixture)
# 
# write_csv(bull_frog_nmixture_3_visits, here("rv_bullfrog.csv"))

library(unmarked)

# first survey period
bull_frog_nmixture_3_visits <- read.csv("rv_bullfrog.csv")


bf_umf_3_1 <- unmarkedFramePCount(y = first_bull_frog_nmixture_3_visits)

summary(bf_umf_3_1)

fm1_bf3_1 <- pcount( ~ 1 ~ 1, 
               data = bf_umf_3_1, 
               engine = "R",
               control=list(trace=T, REPORT=5))

summary(fm1_bf3_1)


backTransform(fm1_bf3_1, "state")

backTransform(fm1_bf3_1, type = "det")


# second survye period
bf_umf_3_2 <- unmarkedFramePCount(y = second_bull_frog_unmarked_nxmixture)

summary(bf_umf_3_2)

fm1_bf3_2 <- pcount( ~ 1 ~ 1, 
               data = bf_umf_3_2, 
               engine = "R",
               control=list(trace=T, REPORT=5))

summary(fm1_bf3_2)


backTransform(fm1_bf3_2, "state")

backTransform(fm1_bf3_2, type = "det")


  


```

```{r}

R <- 4 # number of sites
J <- 3 # number of visits
y <- matrix(c(
   1,2,0,
   0,0,0,
   1,1,1,
   2,2,1), nrow=R, ncol=J, byrow=TRUE)
y
```

